---
title: "p8105_hw3_wx2233"
author: "Weijia Xiong"
date: "10/3/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
library(skimr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 10, 
  fig.height = 8,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + 
            theme(legend.position = "bottom", 
                  title = element_text(size = 20),
                  axis.title = element_text(size = 15)
                  )
)
```

## Problem 1 
```{r}
data("instacart")
```

The data has 1384617 obs. of  15 variables:

* order_id: the id of order 
* user_id: the id of customer 
* eval_set: which evaluation set this order belongs in
* order_number: the order sequence number for this user (1 = first, n = nth)
* order_dow: the day of the week the order was placed on
* order_hour_of_day: the hour of the day the order was placed on
* days_since_prior: days since the last order, capped at 30 (NAs for order_number = 1)
* product_id: the id of product
* product_name: name of the product
* aisle_id: the id of aisle
* aisle: the name of the aisle
* department_id: department identifier
* department: the name of the department
* add_to_cart_order: order in which each product was added to cart
* reordered: whether this product has been ordered by this user in the past(1 = yes, 0 = no)

  
Example of obs in first row:     
The product name is Bulgarian Yogurt, it belongs to yogurt aisle(id = 120) and dairy eggs department(id = 16).The days since prior order is 9 and the hour of the day the order was placed on is 10. The day of the week the order was placed on is Thusday and the order sequence number is 4. And the evaluation set is train. The user id is 112108.
```{r}
instacart[1,]
```


Structure: each row is an observation, each column is a variable.
```{r}
str(instacart)
```



### a) 
```{r}
aisles_data = instacart %>%
  group_by(aisle) %>%
  summarize(count = n()) %>% 
  arrange(desc(count))
aisles_data
number_of_aisles = nrow(aisles_data)
```
There are 134 aisles. Fresh vegetables are the most items orderd from.

### b)

```{r}
aisles_data %>%  
  filter(count >= 10000) %>%  #more than 10000 items ordered
  ggplot(aes(x = reorder(aisle,desc(count)), y = count,fill = count)) +  
  geom_bar(stat = 'identity') +
  labs(
    title = "The Number of Items ordered in Each Aisle",
    y = "The number of Item Orders",
    x = 'Aisles') +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
 
```

The plot shows the number of items ordered in each aisle. From this plot we can also see that fresh vegetables is the aisle which has most items ordered.

### c)  
```{r}
bake = instacart %>% 
  filter(aisle == c("baking ingredients")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(
    count = n()
  ) %>%
  arrange(desc(count)) 
bake = bake[c(1:3),]

dog = instacart %>% 
  filter(aisle == c("dog food care")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(
    count = n()
  ) %>%
  arrange(desc(count)) 
dog = dog[c(1:3),]

veg_fruit = instacart %>% 
  filter(aisle == c("packaged vegetables fruits")) %>% 
  group_by(aisle, product_name) %>% 
  summarize(
    count = n()
  ) %>%
  arrange(desc(count)) 
veg_fruit = veg_fruit[c(1:3),]

bind_rows(bake,dog,veg_fruit) %>% 
  knitr::kable()
```

The most popular item of baking ingredients	is Light Brown Sugar, which has bought been 499 times. The most popular item from dog food care is	Snack Sticks Chicken & Rice Recipe Dog Treats, which has been bought 30 times. And the most popular item from packaged vegetables fruits is	Organic Baby Spinach, which has been bought 9784 times.

### d) Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) %>%
  mutate(
    order_dow = factor(order_dow, labels  = 
                             c("Sunday","Monday","Tuesday","Wednesday",
                              "Thursday","Friday","Saturday"))
    ) %>% 
  group_by(product_name,order_dow) %>% 
  summarize(
    mean_hour = round(mean(order_hour_of_day),2)
  ) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>%
  knitr::kable()

```

From this table we find that the pink lady apple has least orders on Monday and most orders on Wednesday. Coffee Ice Cream has least orders on Friday but most on Tuesday. 




## Problem 2


### Tidy data
```{r}
data("brfss_smart2010")
brfss = brfss_smart2010

brfss = brfss %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health", response %in% c("Excellent","Very good","Good","Fair","Poor")) %>% 
  mutate(
    response = factor(response,levels = c("Poor","Fair","Good","Very good","Excellent"))
  ) %>% 
  rename(state = locationabbr)

brfss
```

### a) 
```{r}

brfss %>% 
  filter(year == 2002 | year == 2010) %>% 
  group_by(year, state) %>% 
  distinct(locationdesc) %>% 
  summarize(
    count = n()
  ) %>% 
  filter(count >= 7) %>% 
  knitr::kable()


```

2002: CT,FL,MA,NC,NJ,PA    
2010: CA,CO,FL,MA,MD,NC,NE,NJ,NY,OH,PA,SC,TX,WA


### b) 
```{r}
ext_avg_data_value = brfss %>% 
  filter(response == "Excellent") %>% 
  select(year,state,data_value) %>% 
  group_by(year,state) %>% 
  summarize(
    average_value = mean(data_value,na.rm = TRUE)
  )

ext_avg_data_value
```


```{r}
ext_avg_data_value %>% 
  ggplot(aes(x = year, y = average_value)) +
  geom_line(aes(group = state,color = state)) +
  theme(
    legend.position = "right"
  ) +
  labs(
    y = "Average Value",
    x = "Year",
    title = "Average value over time within a state"
  )


```

The average data value fluctuates by years for each state. And the average values are different among each state.


### c) 
```{r}
NY_data_value = brfss %>%
  filter(year == 2006 | year == 2010,state == "NY") %>%
  select(year,state,locationdesc,response,data_value)


#boxplot: distribtion among locations. 
NY_data_value %>%
ggplot(aes(x = response, y = data_value)) +
  geom_boxplot() +
  labs(
    title = " Data value of NY"
  ) +
  facet_grid(. ~ year) 

#bar plot: to see the difference of each locationdesc
NY_data_value %>%
ggplot(aes(x = response, y = data_value,fill = locationdesc)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_line(aes(group = locationdesc, color = locationdesc)) +
  geom_point(size = 1.5, alpha = 0.5) +
  labs(
    title = " Data value of NY"
  ) +
  facet_grid(. ~ year)
```

We can see that Only 6 counties' data are collected in 2006 while 9 counties in 2010. The data value of "Poor" response is lowest in both two years and the "Good" and "Very good" response are higher in both two years. 

## Problem 3

### a) 
```{r}
accel = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    weekday_weekend = recode(day,
                             "Monday" = "weekday",
                             "Tuesday" = "weekday",
                             "Wednesday" = "weekday",
                             "Thursday" = "weekday",
                             "Friday" = "weekday",
                             "Saturday" = "weekend",
                             "Sunday" = "weekend"),
    weekday_weekend = factor(weekday_weekend),
    day = factor(day,
                 levels = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))
  ) %>% 
  select(week,weekday_weekend,everything()) %>% 
  pivot_longer(
    starts_with("activity_"),
    names_to = "activity_minute", 
    names_prefix = "activity_",
    values_to = "counts") %>% 
  mutate(
    activity_minute = as.numeric(activity_minute)
  )

accel
```

The data has 50400 obs. of 6 variables:

* week: which week
* weekday_weekend: whether weekday or weekend
* day_id: id of the day
* day: factor variable shows which day of a week
* activity_minute: the minute of a 24-hour day
* count: the activity counts for each minute of a 24-hour day starting at midnight.

### b) 
```{r}
accel_new =
accel %>% 
  arrange(week,day) %>%
  mutate(
    day_index = rep(c(1:35),each = 1440)  #add new day_id 1:35
  ) %>% 
  select(week,day,day_index,everything())

eachday_total = accel_new %>% 
  group_by(week, day,day_index) %>% 
  summarize(
    total = sum(counts,na.rm = TRUE)
  ) %>% 
  select(day_index,everything())

knitr::kable(eachday_total)
```

From the table I cannot see the trends clearly so I draw a plot for 35 days.
```{r}
eachday_total %>% 
  ggplot(aes(x = day_index, y = total )) +
  geom_line() +
  scale_x_continuous(breaks = seq(0, 35, 2))
```

The activity counts have upper and lower fluctuation. Decreasing always occurs after increasing of the day before. And the activity counts dramaticly decrease on day 27 and 34, both are Saturday. 


### c) 

```{r}
accel_new %>%
  ggplot(aes(x = activity_minute, y = counts, color = day)) +
  geom_line(stat = "identity") +
  scale_x_continuous(breaks = seq(0,1440,60),label = c(0:24)) +
  labs(
    y = "activity for each day",
    x = "24 hours",
    title = "The 24-hour Activity Time Courses for Each Day"
  )
  
```

We can find that the activity counts during the morning 7am - 9 am and during the evening around 7pm - 9pm are higher. The activity counts are lower during midnight and early morning. But I cannot disguish each day clealy, so I draw another plot to gain more information.

```{r}
eachday =
accel_new %>% 
  mutate(
    act_min = activity_minute + 
      1440 * (day_index - 1))

eachday %>% 
  ggplot() +
  geom_linerange(aes(x = act_min, ymax = counts, ymin = 0, color = day)) +
  scale_x_continuous(breaks = seq(0,28*1440,1440*7),label = c(1:5)) +
  labs(
    y = "activity for each day",
    x = "week",
    title = "The 24-hour Activity Time Courses for Each Day"
  )
```

The activity counts are low on from Monday to Thursday of the first week.
Then the activity counts fluctuates for each day, with around 2 peaks for morning and evening. And for Saturday in last two weeks, both activity counts of the day dramatically decrease to 1440.
